---
title: "lab_12"
author: "Derek Willis"
date: "2025-01-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## You will need

-   tidytext and our usual libraries

## Load libraries and establish settings

**Task** Create a codeblock and load appropriate packages and settings for this lab.

```{r}
library(tidyverse)
library(tidytext)
library(janitor)
library(lubridate)
library(rvest)
```

## Questions

**Q1.** You've been assigned to report a story about the leading reasons that Maryland attorneys get sanctioned by the state for misconduct. The state [publishes lists of sanctions](https://www.courts.state.md.us/attygrievance/sanctions) that contain a short text description about the situation. Load the CSV file in the data folder containing records from fiscal year 2011 onwards. Make a list of unique words from the text column, then following the example in the pre_lab, remove common "stop words" from that list and create a list of the top 10 words containing the percentage of occurrences each word represents. What's the leading word in that answer and, broadly, what do you think the top 10 words describe?

**A1.** The leading word is "failing", with a frequency of 1794 or about 4.5 percent. The top 10 words broadly describe attorneys being sanctioned for failing their clients financially, likely through mismanaging trusts (5th most common word) among other things.

```{r}
library(readr)
attorney_sanctions <- read_csv("data/md_attorney_sanctions.csv")

unique_words <- attorney_sanctions |> select(text) |>
  unnest_tokens(word, text)

data("stop_words")

unique_words |>
  anti_join(stop_words) |>
  group_by(word) |>
  tally(sort=TRUE) |>
  mutate(percent = (n/sum(n))*100) |>
  top_n(50)


```

**Q2.** Let's move beyond single words to phrases. Make a list of the top 10 three-word phrases, called trigrams, based on the example from the pre_lab (you'll need to modify the example code to do this). What's the top trigram and how often does it appear? What does that phrase mean in legal terms?

**A2.** The top trigram is "attorney trust account," which appears 393 times. In legal terms, that is a bank account that attorneys are required by law to hold money in that belongs to their clients.

```{r}
attorney_sanctions |>
  unnest_tokens(trigram, text, token = "ngrams", n = 3) |>
  drop_na() |>
  separate(trigram, c("word1", "word2", "word3"), sep = " ") |>
  filter(!word1 %in% stop_words$word) |>
  filter(!word2 %in% stop_words$word) |>
  filter(!word3 %in% stop_words$word) |>
  mutate(trigram = paste(word1, word2, word3, sep = " ")) |>
  group_by(trigram) |>
  tally(sort = TRUE) |>
  mutate(percent = (n / sum(n)) * 100) |>
  top_n(10)
```

**Q3.** Let's drop back down to more traditional text analysis - take the top trigram from Q2 and write code to see how many times it occurs in the text column in each fiscal year. What do you think the answer produced by your code suggests? What else could you do to try and clarify the most important reasons attorneys get sanctioned?

**A3.** Attorney trust account sanctions occurred most often during the 2021 fiscal year, at 37 times. The code suggests that sanctions regarding attorney trust accounts steadily rose from 2011 through 2021, where it reached a peak, then slightly decreased in 2022 before spiking down in 2023 and 2024. To clarify the reasons for other sanctions, we could possibly look at bigrams or single words rather than trigrams to cast a wider net, since not every sanction description will use the exact same phrase.

```{r}
attorney_sanctions |>
  filter(str_detect(text, "attorney trust account")) |>
  group_by(fiscal_year) |>
  summarise(count = n()) |>
  print(n = 14)
```
