---
title: "lab_08"
author: "Derek Willis"
date: "2025-01-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings

**Task** Create a codeblock and load appropriate packages and settings for this lab. We'll be making some charts, working with dates and retrieving Census data.

```{r}
library(tidyverse)
library(janitor)
library(lubridate)
library(tidycensus)
```

## Load and modify data

**Task** Create a codeblock and load the following data from the data folder:

-   Prince George's County 911 Overdose Calls

```{r}
pg_calls <- read_csv ("data/prince_georges_2022_overdoses.csv") |> clean_names()

pg_calls <- pg_calls |> mutate(
  date = date(datetime),
  month = month(datetime, label=TRUE),
  week = week(datetime)
)
```

You will need to create columns for the date, week and month based on the existing `datetime` column.

## Questions

**Q1.** Which month saw the greatest percentage of total calls? Create a dataframe that calculates the percentage of all calls that each month's calls represents. Make a bar chart from that dataframe. Your bar chart must have:

-   A clear title that states the main idea/finding
-   Good labels for the x & y axis and a caption for the source, which is Prince George's County EMS.
-   Readable bars - the values shouldn't be overlapping

Compare the results here to those from the pre_lab_08 bar chart - are there major differences in the months with the highest and lowest figures? Describe that below.

**A1.** I would say there's a major difference in the month of December, where it takes up 10% of calls in this bar chart but substantially less in the pre-lab chart. Additionally, while August is the most prominent month for overdose calls in the other chart, it's one of the least in this one.

```{r}
pg_calls_month <- pg_calls |> 
  group_by(month) |> 
  summarize(total_calls = n()) |> 
  mutate(pct = (total_calls/sum(total_calls)*100))

pg_calls_month |>
  ggplot() +
  geom_bar(aes(x=month, weight=pct)) +
  labs(
    title = 'One in 10 overdose calls in PG County take place in December',
    y = 'Percentage of calls',
    x = 'Months',
    caption = 'Source: PG County EMS'
  )
```

**Q2.** Let's visualize this data on a weekly basis using a line chart. As in Q1, generate a dataframe with the total number of calls for each week, and then create a line chart to show the distribution of calls over time. Your line chart must have:

-   A clear title that states the main idea/finding
-   Good labels for the x & y axis and a caption for the source, which is Prince George's County EMS.
-   Readable labels

Describe the pattern of results; you may want to look at the data to dig into particular outliers.

**A2.** The number of calls per week doesn't fluctuate too heavily, generally staying in the 20s and 30s range. The outliers come in January, where the number of calls is below 20 for 3 of the month's 4 weeks, and in Week 28, where it quickly peaks at 41 calls then dips back down.

```{r}
pg_calls_week <- pg_calls |>
  mutate(week = week(date)) |>
  group_by(week) |>
  summarize(total_calls = n())
  
pg_calls_week |>
  ggplot() +
  geom_line(aes(x = week, y = total_calls)) +
  labs(
    title = 'PG County overdose calls peak in Week 28',
    x = 'Week of year',
    y = 'Total calls',
    caption = 'Source: PG County EMS'
  )
```

**Q3.** A scatterplot is a type of chart that helps us see relationships between two variables. One variable goes on the x axis, the other on the y axis. For each row/observation in our data, a scatterplot puts a circle (or a "point") where the two variables intersect on a grid.

Statisticians use scatterplots to show graphically whether one variable is correlated -- related, in a statistical sense -- with another variable. A classic example is the [relationship between ice cream sales and temperature](https://www.mathsisfun.com/data/scatter-xy-plots.html). The scatterplot below -- press play to load the image -- shows that relationship, that an increase in temperature is associated with an increase in ice cream sales. When it's 12C, sales are 200 dollars, and when it's hotter, 25C, sales are 600 dollars.

```{r}
knitr::include_graphics("https://www.mathsisfun.com/data/images/scatter-ice-cream1.svg")
```

We're going to use a scatterplot a little differently, to get a visual sense of two key variables:

Our question is: does the median income in a zip code have any relationship to the number of overdose 911 calls in that zip code?

To answer this question, do the following:

1.  Generate a dataframe with the number of 911 calls for each zip code.
2.  Get data from the Census Bureau showing median household income for Maryland zip codes.
3.  Join those two dataframes on their zip code columns, starting with the 911 calls dataframe.
4.  Make a scatterplot showing the total calls and median income. I didn't show you how to do this, so look it up! Googling "ggplot scatterplot" is a good start.
5.  Give it an appropriate title, source, and x and y axis titles.
6.  Add a label for each point that shows the zip code using geom_text() - see some examples of its use at <https://ggplot2.tidyverse.org/reference/geom_text.html#ref-examples>. Try to make the names as easy to read as possible by avoiding overlap.
7.  In the answer space below, describe what you see and answer the questions posed above. In a general sense, what do you think this means? Feel free to consider the actual raw values: how would you report out the main point(s) of this chart?

**A3.** It seems like the median income in a zip code does have a relationship with the number of overdose calls in that zip code. The higher the median household income, the less income calls there are generally. The highest number of overdose calls come in the zip code 20743, which has an average household income under \$75,000, while the fewest comes from a group of zip codes with average household incomes above \$150,000.

```{r}
pg_calls_zip <- pg_calls |>
  group_by(zipcode) |>
  summarize(total_calls = n())

md_income <- get_acs(
  geography = "zcta",
  variables = "B19013_001",
  year = 2022
)

pg_calls_income <- pg_calls_zip |>
  left_join(md_income, by = c("zipcode" = "GEOID"))

pg_calls_income |>
  ggplot(aes(x = estimate, y = total_calls, label = zipcode)) +
  geom_point() +
  geom_text(check_overlap = TRUE, size = 3, vjust = -0.5) +
  labs(
    title = 'Higher-income zip codes see fewer overdose calls in PG County',
    x = 'Median household income',
    y = '911 overdose calls',
    caption = 'Sources: PG County EMS, U.S. Census Bureau ACS'
  )
```
