---
title: "Reverse Engineering Project"
author: "Student names here"
date: "Date here"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

In this notebook, we are reverse engineering the story, [STORY NAME](STORY%20URL)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
library(lubridate)
library(janitor)
library(tidyverse)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data.

```{r}
# Load required data

# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

# Clean required data and prepare for analysis if needed.

sports_political_donations <- read_csv("data/sports-political-donations.csv") |> clean_names()


```

## Sentences to Engineer

In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

-   **Sentence text**: [Sports owners' contributions to federal elections since 2015: \$47 million (\$10m to republicans, \$1.9 million to Dems in 2020 as of publishing)]
-   **Analysis summary**: I was able to confirm the finding for these numbers. I just did a simple filter for 2020 election, grouped by party and summarized the total amount for each of those parties. The most difficult part of this analysis was figuring out how to change the "amount" column from a character to a number. Me and my partners weren't sure how to do that, so we prompted ChatGPT with that question and it gave us the top part of the below code.

```{r}
# Put code to reverse engineer sentence here
clean_numeric <- function(x) {
  as.numeric(gsub("[$,]", "", x))
}

sports_political_donations$amount <- clean_numeric(sports_political_donations$amount)

sports_political_donations |>
    filter(election_year == 2020) |>
    group_by(party) |>
    summarise(total = sum(amount))

# Display results of code below this codeblock
```

### Sentence 2

-   **Sentence text**: [Paste in sentence to engineer here]
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 3

-   **Sentence text**: [Paste in sentence to engineer here]
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 4

-   **Sentence text**: [Paste in sentence to engineer here]
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 5

-   **Sentence text**: [More than 63 percent of the \$47 million in the database of nearly 3,000 contributions comes from just 108 donations split among these 40 owners --- and 79 percent of that money went to Republican causes.]
-   **Analysis summary**: [Write up two to three sentences describing the results of your analysis. Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here


# Display results of code below this codeblock

```

-30-
