---
title: "week7_recap"
author: "Daniel Trielli"
date: "2023-10-04"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Turn off scientific notation
options(scipen=999)
```

## Loading the packages

Run the codeblock below to load the packages we will need for this recap

```{r}
library(tidyverse)
library(lubridate)
library(janitor)
```

## Load Data

Run the codeblock below to load the data.

```{r}
earthquakes <- read_csv('https://earthquake.usgs.gov/earthquakes/feed/v1.0/summary/all_month.csv')

#Setting time column as datetime
earthquakes <- earthquakes |> mutate(time = as_datetime(time))
```

#### Answer the questions below

Most questions have a code block and a space for an answer below. Write the code you think is necessary and, in the answer space, write out what you did and what was the result.

------------------------------------------------------------------------

#### **Q1** Do you notice anything different with this read_csv()? Particularly in the file name?

**A1: The file is different, it's a straight URL link rather than the usual format starting with "data/" etc.**

------------------------------------------------------------------------

#### **Q2** How many records there are there in this dataset? What do they mean and what useful information we can gather from it, looking at the columns?

**A2: There are 10,754 observations. It's measuring past earthquakes around the world. Some of the key columns include latitude, longitude, depth, magnitude, location and location source.**

------------------------------------------------------------------------

#### **Q3** How do I reorganize this list to see the ones that are the deepest first? What is the depth that shows up for the deepest one, and its magnitude?

```{r}
earthquakes |>
  arrange(desc(depth))

```

**A3: The deepest earthquake is 750. There are two with that depth, one's magnitude is N/A and the others is 3.6.**

------------------------------------------------------------------------

#### **Q4** I just want to see the earthquakes with a magnitude larger than 6. How do I do that? And how many are there that fit this criteria?

```{r}
earthquakes |>
  filter(mag > 6) |>
  arrange(desc(mag))
```

**A4: There are 12 earthquakes with a magnitude over 6, although I'm not sure why it isn't displaying the last two on the second page of my table.**

------------------------------------------------------------------------

#### **Q5** What about if I want to see earthquakes that have both a magnitude larger than 6 and a depth smaller than 20? How many are there in the data set that fit [both]{.underline} these criteria?

```{r}
earthquakes |>
  filter(mag > 6 & depth < 20)
```

**A5: There are 10 earthquakes that fit both these criteria.**

------------------------------------------------------------------------

#### **Q6** What about if I want to see earthquakes that either have a magnitude larger than 6 and a depth smaller than 20? How many are there in the data set that fit [either]{.underline} these criteria?

```{r}
earthquakes |>
  filter(mag > 6 | depth < 20)
```

**A6: There are 8,549 earthquakes that fit either of these criteria.**

------------------------------------------------------------------------

#### **Q7** I'm interested in finding earthquakes that took place in Alaska. Which column should I look at? How do I use it to find all the earthquakes in Alaska? How many earthquakes took place there?

```{r}
earthquakes |>
  filter(str_detect(place, 'Alaska'))
```

**A7: 3,568 earthquakes took place in Alaska.**

------------------------------------------------------------------------

#### **Q8** I notice that there is a column called 'type', that seems to have different kinds of tremors. What is the best way to find what are all the possible types of tremors, and counting how many of each there are in this data set? What are the first two most common types of tremors in this data set?

```{r}
earthquakes |>
  count(type)
```

**A8: The two most commons tremors are earthquake (10,582) and quarry blast (105).**

------------------------------------------------------------------------

#### **Q9** How do I calculate the average depth of each type of tremor? What is the average depth of the earthquake type in this data set? Is there anything that seems unusual or surprising in this finding?

```{r}
earthquakes |>
  group_by(type) |>
  summarize(avg_depth = mean(depth))
```

**A9: The average depth of an earthquake in this set is 22.79.**

------------------------------------------------------------------------

#### **Q10** I'm interested, in the future, to see the hours in which earthquakes happen. How can I extract the hour from the time column?

```{r}
earthquakes |>
  mutate(hour = hour(time))
```

**A10: This added a new column with the hour of each earthquake in military time.**

------------------------------------------------------------------------

#### **Q11** I want to make a note of all the records in this data set that I consider serious. For my purposes, I'm thinking that all tremors that have a magnitude that is larger than 3 are serious. How do I automatically create a new column making that distinction?

```{r}
earthquakes |>
  mutate(serious = case_when(
    mag > 3 ~ "serious",
    mag <= 3 ~ "not serious"))
```

**A11: This creates a column called serious that lists each earthquake as serious or not serious based on its magnitude.**

------------------------------------------------------------------------

#### **Q12** I have no idea how earthquakes work and I'm interested in seeing if there is a particular time of day in which serious earthquakes happen. How can I see that condensed in a table with all the hours in a day and all the serious earthquakes in each hour? What is the hour with fewer serious earthquakes and the one with the most serious earthquakes?

```{r}
earthquakes |>
  filter(mag > 3) |>
  mutate(hour = hour(time)) |>
  count(hour) |>
  arrange(hour)
```

**A12**: 2:00 is the hour with the fewest serious earthquakes (36), while 6:00 and 17:00 are tied for the most (73).

------------------------------------------------------------------------

#### EXTRA Q What other questions are you interested in and how would you ask them in R?

**Q: How many serious earthquakes happen per day on average?**

```{r}
earthquakes |>
  filter(mag > 3) |>
  mutate(date = date(time)) |>
  count(date) |>
  summarize(average = mean(n))
```

**Q; What percent of tremors in the dataset are earthquakes?**

```{r}
earthquakes |>
  count(type) |>
  mutate(percentage = n / sum(n))
```

**Q: What's the northernmost earthquake that occurred in this dataset?**

```{r}
earthquakes |>
  arrange(desc(latitude))
```
